---
title: "Estimation of site index for YSM samples"
output: github_document
author: Hyeyoung Woo
date: 18 March, 2026
---

```{r setup, echo=FALSE, warning=FALSE}
suppressMessages(library(data.table))
suppressMessages(library(dplyr))
#suppressMessages(library(ggplot2))
suppressMessages(library(tidyverse))
suppressMessages(library(openxlsx))
suppressMessages(library(htmlwidgets))

comp_table <- read.table(text = "
source target value IDsource IDtarget
CONV_EQN CONV_EQN 1275 0 6
CONV_EQN_SUB CONV_EQN 43 1 6
DEFAULT_SI15 CONV_EQN 4 2 6
GROUND_PLOT CONV_EQN 1 3 6
LEAD_SPECIES CONV_EQN 163 4 6
PSPL_SI CONV_EQN 261 5 6
CONV_EQN CONV_EQN_SUB 0 0 7
CONV_EQN_SUB CONV_EQN_SUB 7 1 7
DEFAULT_SI15 CONV_EQN_SUB 0 2 7
GROUND_PLOT CONV_EQN_SUB 0 3 7
LEAD_SPECIES CONV_EQN_SUB 13 4 7
PSPL_SI CONV_EQN_SUB 2 5 7
CONV_EQN DEFAULT15 0 0 8
CONV_EQN_SUB DEFAULT15 0 1 8
DEFAULT_SI15 DEFAULT15 36 2 8
GROUND_PLOT DEFAULT15 0 3 8
LEAD_SPECIES DEFAULT15 2 4 8
PSPL_SI DEFAULT15 0 5 8
CONV_EQN GROUND_SI_SUIT_NULL 0 0 9
CONV_EQN_SUB GROUND_SI_SUIT_NULL 0 1 9
DEFAULT_SI15 GROUND_SI_SUIT_NULL 16 2 9
GROUND_PLOT GROUND_SI_SUIT_NULL 514 3 9
LEAD_SPECIES GROUND_SI_SUIT_NULL 9 4 9
PSPL_SI GROUND_SI_SUIT_NULL 237 5 9
CONV_EQN GROUND_SI_SUIT_NULL_CONV_EQN 271 0 10
CONV_EQN_SUB GROUND_SI_SUIT_NULL_CONV_EQN 0 1 10
DEFAULT_SI15 GROUND_SI_SUIT_NULL_CONV_EQN 1 2 10
GROUND_PLOT GROUND_SI_SUIT_NULL_CONV_EQN 3 3 10
LEAD_SPECIES GROUND_SI_SUIT_NULL_CONV_EQN 81 4 10
PSPL_SI GROUND_SI_SUIT_NULL_CONV_EQN 236 5 10
CONV_EQN GROUND_TLSO 0 0 11
CONV_EQN_SUB GROUND_TLSO 0 1 11
DEFAULT_SI15 GROUND_TLSO 0 2 11
GROUND_PLOT GROUND_TLSO 1928 3 11
LEAD_SPECIES GROUND_TLSO 0 4 11
PSPL_SI GROUND_TLSO 0 5 11
CONV_EQN GROUND_TLSOX 0 0 12
CONV_EQN_SUB GROUND_TLSOX 0 1 12
DEFAULT_SI15 GROUND_TLSOX 0 2 12
GROUND_PLOT GROUND_TLSOX 16 3 12
LEAD_SPECIES GROUND_TLSOX 0 4 12
PSPL_SI GROUND_TLSOX 0 5 12
CONV_EQN LEAD_SPECIES 0 0 13
CONV_EQN_SUB LEAD_SPECIES 0 1 13
DEFAULT_SI15 LEAD_SPECIES 4 2 13
GROUND_PLOT LEAD_SPECIES 0 3 13
LEAD_SPECIES LEAD_SPECIES 259 4 13
PSPL_SI LEAD_SPECIES 0 5 13
CONV_EQN PSPL_SI 0 0 14
CONV_EQN_SUB PSPL_SI 0 1 14
DEFAULT_SI15 PSPL_SI 0 2 14
GROUND_PLOT PSPL_SI 2 3 14
LEAD_SPECIES PSPL_SI 0 4 14
PSPL_SI PSPL_SI 450 5 14
CONV_EQN PSPL_SI_SUB 0 0 15
CONV_EQN_SUB PSPL_SI_SUB 0 1 15
DEFAULT_SI15 PSPL_SI_SUB 3 2 15
GROUND_PLOT PSPL_SI_SUB 0 3 15
LEAD_SPECIES PSPL_SI_SUB 25 4 15
PSPL_SI PSPL_SI_SUB 53 5 15
CONV_EQN PSPL_SP_CONV_EQN 0 0 16
CONV_EQN_SUB PSPL_SP_CONV_EQN 0 1 16
DEFAULT_SI15 PSPL_SP_CONV_EQN 3 2 16
GROUND_PLOT PSPL_SP_CONV_EQN 0 3 16
LEAD_SPECIES PSPL_SP_CONV_EQN 3 4 16
PSPL_SI PSPL_SP_CONV_EQN 0 5 16
", header = TRUE, stringsAsFactors = FALSE)
```


Estimation of site index for YSM samples involves integrating multiple data sources and methods. The basic strategy is as follows:

*	Use ground-measured site index where available  
*	Apply site index conversion equations as needed  
*	Incorporate PSPL site index  
*	Use substitutes when necessary, including:  
    -  Average of known values  
    -  Leading species  
    -  Default values  

Following these steps, the source of site index is classified according to the hierarchy below.  


**1.	GROUND_TLSO**  
ISMC-compiled site index based on T, L, S, and O trees  

*  Use only records with `SUIT_SI = 'Y'`  
*  Exclude trees that should not be considered suitable as site trees `SUIT_SI = 'N'`;  
    -  if `AGE_BH > 0` and `AGE_BH < 5` or `AGE_BH > 150`   
    -  if `HEIGHT > 0` and `HEIGHT <= 1.5`   


**2.	GROUND_TLSOX**  
ISMC-compiled site index, based on T, L, S, O, and X trees   

*  Approximately 50–60 trees with `TH_TREE = 'X'` were sampled in YSM plots  
*  These represent _"Additional X trees may also be selected in a non-standard manner at the discretion of the field crew"_.  
*  They are intended to represent age, height, and SI characteristics of species that may be:  
    -  neither leading nor secondary species,  
    -  outside the plot, or  
    -  lower in the diameter ranking.  


**3.	CONV_EQN**  
Site index derived using site index conversion equations.  

*  Use the conversion equations available in the [FAIB PSPL repository](https://github.com/bcgov/FAIB_PSPL/tree/main/05_site_index_conversions).  
*  These equations were converted from *SIndex*, with additional species substitutions incorporated and approved by Gord Nigh.    

```{r, echo=F}
spec_match <- c("BA" = "BG", "BA" = "BL", "BL" = "BA", "BL" = "BG", "BG" = "BA",
                  "DR" = "FD", "PY" = "FD", "SS" = "PW", "LT" = "LW", "LW" = "LT", 
                  "HM" = "HW", "HW" = "HM", "PA" = "PL", "PL" = "PA",
                  "SW" = "SX", "SW" = "SE", "SE" = "SW")
  
knitr::kable(data.table(t(spec_match)),format = "markdown")
```


**4.	GROUND_SI_SUIT_NULL**  
Site index derived from the compiled output when suitability is not assessed under earlier standards.  

*  In ISMC-compiled site tree, `SUIT_SI` is `NULL` and the trees are `SUIT_HT = 'Y'` and `SUIT_AGE = 'Y'`, and `CROWN_CLASS_CODE = 'D'` or `'C'` and no broken top   


**5.	GROUND_SI_SUIT_NULL_CONV_EQN**  
Apply the site index conversion equations (described in 3.) to GROUND_SI_SUIT_NULL values  


**6.	PSPL_SP**  
PSPL-derived site index  

*  [PSPL version 7](https://github.com/bcgov/FAIB_PSPL/tree/main/03_pspl_mean_values) 
*  For the species where PSPL is not available, species substitutions are incorporated (**PSPL_SI_SUB**).  


**7.	PSPL_SP_CONV_EQN**  
Apply site index conversion equations to PSPL_SP  


**8.	LEAD_SPECIES**  
Use the site index of the leading species when none of the above sources are available for the species  


**9.	DEFAULT_SI15**  
Use a default site index value of 15 when none of the above sources are available



When multiple site index values are available for the same site and species due to repeated visits and measurements, the following procedure is applied.  

  
*  For each species in each sample visit, compute a site index using the ground-measurement prioritization steps described above.  
*  When multiple visits produce more than one SI estimate for a species, the estimate closest to 50 years at breast height (BH) is prioritized, with preference given to ground-based sources (1-5 above). The selection follows these rules:  
    -  Closeness to age 50 BH is defined as the absolute difference from 50 years BH.  
    -  If two estimates are equally close, the older age is preferred (e.g., if visit 1 = 48 years and visit 2 = 52 years, the 52-year estimate is selected).  
    -  If the estimate closest to age 50 BH has a ground-based source (1-5 above), it is selected.  
    -  If not, the next closest estimate to age 50 BH with a ground-based source (1-5 above) is selected, checking sequentially (third closest, fourth closest, etc.).  
    -  If no ground-based SI exists for that species across any visits, the estimate closest to age 50 BH is used as a last resort (6-9 above).



### Data source 

**Site index conversion equation parameter**   

-  [FAIB PSPL repository](https://github.com/bcgov/FAIB_PSPL/tree/main/05_site_index_conversions)  
-  [Download TIPSY 44 version (incomplete)](//objectstore3.nrs.bcgov/s164/S63016/!Workgrp/Inventory/Compilation/ismc/external_inputs/si_conv_eqns/conv_eqn_tipsy44.xlsx) - IDIR access required  


**PSPL 7 overlay on ground samples**  

-  [Access to folder](//objectstore3.nrs.bcgov/s164/S63016/!Workgrp/Inventory/Compilation/ismc/external_inputs/spatial_overlay/pspl_overlay) - IDIR access required  


**SIndexR**  

-  [Download SIndexR 2.0](//objectstore3.nrs.bcgov/s164/S63016/!Workgrp/Inventory/Site Productivity/SIndexR/SIndexR_2.0.tar.gz) - IDIR access required  


Based on the list above, the site index of all species in the YSM samples is reclassified as below.  

-  Left: Site index sources used in the current TASS projection (20250514 compilation)  
-  Right: Reclassified site index sources  


```{r, echo = F, results='asis', message=FALSE, warning=FALSE, width = 800}
suppressWarnings(library(networkD3))

#colnames(comp_table) <- c("source", "target", "value")
comp_table <- comp_table %>% filter(value != 0)
comp_table$target <- paste(comp_table$target, " ", sep="")

nodes <- data.frame(name=c(as.character(comp_table$source), as.character(comp_table$target)) %>% unique())

comp_table$IDsource=match(comp_table$source, nodes$name)-1
comp_table$IDtarget=match(comp_table$target, nodes$name)-1

ColourScal ='d3.scaleOrdinal() .range(["#FDE725FF","#B4DE2CFF","#6DCD59FF","#35B779FF","#1F9E89FF","#26828EFF","#31688EFF","#3E4A89FF","#482878FF","#440154FF"])'

sankeyNetwork(Links = comp_table, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name",
              sinksRight=F, 
              #colourScale=ColourScal, 
              nodeWidth=40, fontSize=13, nodePadding=20)

```


### Example code

R code and functions used to generate the site index table for YSM samples are archived  [here](https://github.com/hyewoo/ysm_functions/blob/main/siteindextable_functions.r). Note that the functions require data in a specific format.


```{r eval=F}
source("https://raw.githubusercontent.com/hyewoo/ysm_functions/refs/heads/main/siteindextable_functions.r")

library(data.table)
library(dplyr)
library(tidyverse)
library(SIndexR) 

# 1 TLSO site trees
tlsotrees <- select_sitetrees(tree_ysm,"TLSO", suit = "Y")
si_tls0 <- compute_ground_si(tlsotrees,"GROUND_TLSO")

# 2 TLSOX
tlsoxtrees <- select_sitetrees(tree_ysm,"TLSOX", suit = "Y")
si_tls0x <- compute_ground_si(tlsoxtrees,"GROUND_TLSOX")

# merge
si <- plotspc_ysm %>%
  left_join(si_tls0, by = c('MGMT_UNIT', 'SITE_IDENTIFIER', 'VISIT_NUMBER', 'SPECIES', 'BEC_ZONE')) %>%
  left_join(si_tls0x, by = c('MGMT_UNIT', 'SITE_IDENTIFIER', 'VISIT_NUMBER', 'SPECIES', 'BEC_ZONE'), suffix=c("","_x"))

si <- si %>%
  mutate(
    si_grd = coalesce(si_grd, si_grd_x),
    si_src = coalesce(si_src, si_src_x)
  )

# 3 conversion
si <- apply_conv(si,"CONV_EQN")

# 4 relaxed suitability trees
suit_trees <- select_sitetrees(tree_ysm,"TLSOX", suit = "N")
si_suit <- compute_ground_si(suit_trees,"GROUND_SI_SUIT_NULL")

si <- si %>%
  left_join(si_suit, by = c('MGMT_UNIT', 'SITE_IDENTIFIER', 'VISIT_NUMBER', 'SPECIES', 'BEC_ZONE')) %>%
  mutate(
    si_grd = coalesce(si_grd.x, si_grd.y),
    si_src = coalesce(si_src.x, si_src.y)
  )

# 5 conversion again
si <- apply_conv(si,"GROUND_SI_SUIT_NULL_CONV_EQN")

# 6 substitution species
si <- apply_species_substitution(si) 

# 7 PSPL
si <- fill_pspl(si, pspl)

# 8 PSPL conversion
si <- apply_conv(si,"PSPL_SP_CONV_EQN")

# 9 lead species rule
si <- apply_lead_species_rule(si)

# 10 default
si <- apply_default_rule(si)

```
