---
title: "`r title()` Young Stand Monitoring Program"
author: "Forest Analysis and Inventory Branch"
date: "Version 2024-12-19; Created `r Sys.Date()`"
#runtime: shiny
always_allow_html: true
geometry: margin=1cm
output:
  pdf_document:
    keep_tex: true
    latex_engine: xelatex
    toc: true
  html_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.pos = "H", out.extra = "")
knitr::opts_chunk$set(echo = FALSE, fig.align="center")
knitr::opts_chunk$set(fig.height = 4, collapse = TRUE)
knitr::opts_chunk$set(dev = "ragg_png")
set_flextable_defaults(fonts_ignore=TRUE)

htmltopdf <-function(x) {
  if (input$format == "HTML"){
    x
    } else {
      pandoc::pandoc_convert(text = x, from = "html", to = "markdown")
      }
}

colorize <- function(x, color) {
  if (knitr::is_latex_output()) {
    sprintf("\\textcolor{%s}{%s}", color, x)
  } else if (knitr::is_html_output()) {
    sprintf("<span style='color: %s;'>%s</span>", color, x)
  } else x
}
```

<style>
.leaflet {
    margin: auto;
}
</style>


## Overview

`r htmltopdf(overview())`

```{r plotgraph, fig.width=8, echo=FALSE, fig.align='center'}
plotgraph()
```

<br>

```{r flex, echo=FALSE,warning=FALSE, message = FALSE}
flex()
```


## Summary of Key Findings for Existing Young Stands in `r title()` related to Timber Supply

```{r, echo=FALSE, include=FALSE}

projectiontable <- projectiontable()
prjtab_70 <- projectiontable %>% filter(AGE >=70, AGE <100)
  max_row = which.max(abs(prjtab_70$meanvoldiff/prjtab_70$meanvol_tass*100))
  
  maxvoldiff = prjtab_70$percvoldiff[max_row]
  ageatmaxvoldiff = prjtab_70$AGE[max_row]
  Significant = ifelse(prjtab_70$pval[max_row] <0.05, "Yes", "No")
  TSRbias1 = ifelse(prjtab_70$meanvoldiff[max_row] < 0, "Conservative", "Optimistic")
  TSRbias2 = ifelse(Significant == "No", "No", TSRbias1)
  
```


1. The leading species from YSM ground samples is compared to the 
    interpreted leading species in the Vegetation Resources Inventory forest 
    inventory coverage (VRI). The leading species percent agreement is: 
    **`r colorize(correct_ls(), "red")`** `r htmltopdf("<span style='color:red;'>(% agreement)</span>")`
    
2. Species composition from YSM ground samples is compared against 
    TSR inputs (ie., regeneration assumptions used in modeling existing stands 
    in Timber Supply Review [TSR]). The overall species composition overlap is: 
    **`r colorize(round(percoverlap(), 0), "red")`** <span style='color:red;'>(% overlap)</span>
    
3. The Provincial Site Productivity Layer (PSPL), one of the TSR 
    inputs used for modeling existing managed stands, is assessed for bias 
    from YSM ground based site index data. If significant, site index bias 
    (in percent) is listed by species, where a positive percent is an 
    under-estimate in the PSPL, and a negative percent is an over-estimate: 
    **`r colorize(si_bias(), "red")`** 
    
4. YSM sample measurements include conifer and deciduous tree species. 
    The deciduous proportion in the YSM samples (% of total volume) is: 
    **`r colorize(decid_vol(), "red")`** `r colorize("(% of m^3^/ha)", "red")`
    
5. YSM sample measurements include separate tracking of both managed 
    vs. residual cohorts. The residual proportion in the YSM samples 
    (% of total volume) is: 
    **`r colorize(round(fig6_sum(), 0), "red")`** `r colorize("(% of m^3^/ha)", "red")`

6. The periodic annual increment (PAI) of TSR yield tables are compared 
    against re-measured YSM samples over the same remeasurement period, to test 
    if TSR projections are significantly different from YSM growth rates: 
    **`r colorize(test1_comment(), "red")`**
    
7. The PAI of YSM TASS projections are compared against re-measured 
    YSM samples over the same remeasurement period, to test if TASS projections 
    are significantly different from YSM growth rates: 
    **`r colorize(test2_comment(), "red")`**
              
8. For YSM measurements since 2017, the impact from stem rusts can be 
    directly modeled in TASS using GRIM / CRIME. The volume impact of TASS YSM 
    projections by age 100 (in addition to default OAFs) is: 
    **`r colorize(year100_inc(), "red")`** `r colorize("(% of m^3^/ha)", "red")`
    
9. To address forest health risks from all other damage agents, an 
    interim simplistic approach is applied to estimate future impacts of all 
    known forest health agents, and results in an impact by age 100 (in addition 
    to default OAFs) of: 
    **`r colorize(year100_comb(), "red")`** `r colorize("(% of m^3^/ha)", "red")`
     
10. TSR yield tables are compared against YSM TASS projections, to test 
    if TSR assumptions will meet future expectations of young stands by rotation 
    age. Reported outputs include the maximum percent difference of TSR volume 
    relative to YSM (a negative % is where TSR is lower than YSM, positive % is 
    where TSR is greater than YSM); the projected age this occurs at; and a test 
    to determine if the differences between the two is significant. 

   - **`r colorize(paste0("Max % vol diff: ", maxvoldiff), "red")`**
   - **`r colorize(paste0("Age @max vol diff: ", ageatmaxvoldiff), "red")`**
   - **`r colorize(paste0("Significant?: ", Significant), "red")`**
   - **`r colorize(paste0("TSR bias?: ", TSRbias2), "red")`**


## Young Stand Description

`r htmltopdf(ysd())`

```{r, fig.width=5,fig.height=3, echo=FALSE,warning=FALSE, message = FALSE}
summaryflex()
livespplot() + theme_bw(12) +
      theme(
        axis.line = element_line(colour="darkgray"), 
        panel.grid.major.y = element_line(color = 'darkgray'), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        rect = element_blank()
      ) 
```

```{r, fig.width=4,fig.height=3, echo=FALSE,warning=FALSE, message = FALSE}
becplot() + theme_bw(12) + 
      theme(
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_line(colour="darkgray"),
        rect = element_blank()
      )
```

```{r, fig.width=6,fig.height=4, echo=FALSE,warning=FALSE, message = FALSE}
stockplot() + theme_bw(12) +
      theme(
        panel.grid.major.y = element_line(color = 'darkgray'), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        rect = element_blank()
      ) 
```


## Leading Species vs. VRI, & Overall Species Composition vs. TSR Inputs

`r htmltopdf(lstext())`

```{r, fig.height=4,echo=FALSE,warning=FALSE}
lsflex() %>% bg(j = (1:(lsflex()$body$content$ncol/2))*2, 
                     bg = "lightgray", part = "body")
spcomp() + theme_bw(12) +
    theme(
      panel.grid.major.y = element_line(color = 'darkgray'), 
      plot.caption = element_text(hjust=0, size=rel(1.2)),
      legend.position="top",
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      rect = element_blank()
    ) 
```


## Post-Harvest-Regenerated vs. Residual Trees

`r htmltopdf(residtext())`

```{r, fig.height=4,echo=FALSE,warning=FALSE}
residplot() + theme_bw(12) +
    #theme_bw() + 
    theme(
      axis.line = element_line(colour="darkgray"), 
      panel.grid.major.y = element_line(color = 'darkgray'), 
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      rect = element_blank(),
      legend.box.background = element_rect(fill = "white", color = NA),
      legend.position =  "inside", 
      legend.position.inside = c(0.90, 0.90),
      legend.title = element_blank(),
      plot.caption = element_text(hjust=0, size=rel(1.2))
    )  
```


## Site Index vs. Provincial Site Productivity Layer

`r htmltopdf(sitext())`

```{r, fig.height=4,echo=FALSE,warning=FALSE, message = FALSE}
siflex()
```

## Trends in Site Index Estimates over Time

`r htmltopdf(sitrendtext())`

```{r, fig.height=4,echo=FALSE,warning=FALSE, message = FALSE}
siplot() + theme_bw(12) +
    theme(axis.line = element_line(colour="darkgray"), 
          panel.grid.major.y = element_line(color = 'darkgray'), 
          #panel.grid.major.x = element_blank(),
          panel.grid.minor.x = element_blank(),
          rect = element_blank()
    )  
```


## Comparing Current Volumes: TSR Predicted Yield Tables vs. YSM Actual Measurements

`r htmltopdf(curvoltext())`

```{r, fig.height=4,echo=FALSE,warning=FALSE}
ysmtsr() + theme_bw(12) +
    #theme_bw() + 
    theme(
      plot.title = element_text(lineheight = 0.9),
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.major.y = element_line(color = 'darkgray'), 
      rect = element_blank()
    ) 

ysmtsrbias() + theme_bw(12) +
    #theme_bw() + 
    theme(
      legend.position="top",
      #panel.grid.major.x = element_blank(),
      panel.grid.major.y = element_line(color = 'darkgray'), 
      panel.grid.minor.y = element_line(color = 'darkgray'), 
      panel.grid.minor.x = element_blank(),
      rect = element_blank()
    ) 
```


## Test to Compare TSR Total Age vs. YSM Ground Sample Age

`r htmltopdf(agetext())`

::: columns
::: column
```{r, include=T, echo=F,warning=FALSE}
ageflex1()
```
:::

::: column
```{r, include=T, echo=F,warning=FALSE}
ageflex2()
```
:::
:::

```{r, include=T, echo=F, fig.height=2, fig.width=5.5, warning=FALSE, message = FALSE}
ageplot() + theme_bw(12) +
    #theme_bw() + 
    theme(
      axis.ticks = element_blank(),
      panel.grid.major.y = element_blank(),
      panel.grid.minor.x = element_blank(),
      panel.grid.minor.y = element_blank(),
      axis.text.y = element_blank(), axis.ticks.x = element_blank(),
      rect = element_blank()
    ) 
```


## Test to Compare Modeled vs. YSM Re-measured Periodic Annual Increment

`r htmltopdf(paitext())`

::: columns
::: column
```{r, include=T, echo=F,warning=FALSE, message = FALSE}
tsrpai1()
tsrpai2()
```
:::

::: column
```{r, include=T, echo=F, warning=FALSE, message = FALSE}
tasspai1()
tasspai2()
```
:::
:::

```{r, fig.height=2.5, fig.width=6, echo=FALSE, warning=FALSE, message = FALSE}
tassdiff() + theme_bw(12) +
      theme(
        legend.position = "left",
        axis.ticks = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.minor.y = element_blank(),
        axis.text.y = element_blank(), 
        axis.ticks.x = element_blank(),
        rect = element_blank()
      ) 
```


## Quantifying Change in Growth and Mortality

`r htmltopdf(coctext())`

```{r, echo=FALSE,warning=FALSE, message = FALSE}
cocplot() + theme_bw(12) +
      #theme_bw() + 
      theme(
        axis.title.y = element_text(color="steelblue"),
        axis.title.y.right = element_text(colour = "#B4464B"),
        legend.box.background = element_rect(fill = "white", color = NA),
        legend.position = c(0.15, 0.85),
        legend.title = element_blank(),
        axis.line = element_line(colour="darkgray"), 
        panel.grid.major.y = element_line(color = 'darkgray'), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        rect = element_blank()
      ) 
```


## Current Forest Health Incidence

`r htmltopdf(curfhtext())`

```{r, echo=FALSE, fig.height = 4, warning=FALSE, message = FALSE}
curfhplot() + theme_bw(12) +
    #theme_bw() + 
    theme(
      axis.line = element_line(colour="darkgray"), 
      axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
      panel.grid.major.y = element_line(color = 'darkgray'), 
      #panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      rect = element_blank(),
      legend.box.background = element_rect(fill = "white", color = "lightgray"),
      legend.position = "none",
      legend.title = element_blank(),
      plot.caption = element_text(hjust=0, size=rel(1.2))
    )  
```


## Comparing Change in Forest Health Incidence

`r htmltopdf(fhcoctext1())`

```{r, echo=FALSE,warning=FALSE, fig.height = 5,fig.width = 8, message = FALSE}
fhcocplot() + theme_bw(12) +
      theme(
        axis.line = element_line(colour="darkgray"), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        panel.grid.major.y = element_line(color = 'darkgray'), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        rect = element_blank(),
        legend.box.background = element_rect(fill = "white", color = "lightgray"),
        legend.position = c(0.9, 0.8),
        legend.title = element_blank(),
        plot.caption = element_text(hjust=0, size=rel(1.2))
      ) 
```

`r htmltopdf(fhcoctext2())`

```{r, echo=FALSE,warning=FALSE, message = FALSE}
fhcocflex()
```


## Approximating Future Forest Health Risks

`r htmltopdf(fufhtext())`

::: columns
::: column
```{r, echo=FALSE, warning=FALSE, message = FALSE, fig.width=4, fig.height=3}
fufhplot1() + theme_bw(12) +
      theme(
        axis.ticks.x=element_blank(), 
        panel.grid.major.y = element_line(color = 'darkgray'), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        rect = element_blank(),
        legend.title = element_blank()
      )  
```
:::

::: column
```{r, echo=FALSE, warning=FALSE, message = FALSE, fig.width=4, fig.height=3}
fufhplot2() + theme_bw(12) +
      theme(
        axis.ticks.x=element_blank(), 
        panel.grid.major.y = element_line(color = 'darkgray'), 
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        rect = element_blank(),
        legend.title = element_blank()
      )  
```
:::
:::


## Will Existing Young Stands Meet Expectations at Rotation?

`r htmltopdf(projtext())`

```{r, echo=FALSE, warning=FALSE, message = FALSE, fig.height=4, fig.width=9}
projvol() + theme_bw(12) +
    theme(
      axis.line = element_line(colour="darkgray"), 
      panel.grid.major.y = element_line(color = 'darkgray'), 
      panel.grid.major.x = element_blank(),
      panel.grid.minor.x = element_blank(),
      rect = element_blank(),
      legend.box.background = element_rect(fill = "white", color = "lightgray"),
      legend.position = c(0.2, 0.9),
      legend.title = element_blank(),
      plot.caption = element_text(hjust=0, size=rel(0.8))
    )  +
    guides(colour = guide_legend(reverse = TRUE))
```

::: columns
::: column
```{r, echo=FALSE, warning=FALSE, message = FALSE}
stemrusttable()
```
:::

::: column
```{r, echo=FALSE, warning=FALSE, message = FALSE}
cumimptable()
```
:::
:::


## Test to compare YSM TASS projections against TSR Predicted Yield Tables

`r htmltopdf(yieldtext())`

```{r, echo=FALSE, warning=FALSE, message = FALSE}
yieldtable()
```


## Total number of YSM samples by:

::: columns
::: column
```{r, include=T, echo=F,warning=FALSE, message = FALSE}
t1()
t3()
t5()
```
:::

::: column
```{r, include=T, echo=F, warning=FALSE, message = FALSE}
t2()
t4()
```
:::
:::

_** Note the samples defined as 'Excluded' in the TSR yield table current regime 
assignment were dropped from this analysis._


## Tree Species and Damage Agents Recorded from YSM Samples in `r title()`

:::::::::::::: {style="display: grid; grid-template-columns: 1fr 1fr; grid-column-gap: 10px; "}
::: {}
#### Tree Species Codes / Names
```{r, include=T, echo=F, warning=FALSE, message = FALSE}
if (input$format == "HTML"){
    datatable(spcdtab(), rownames= FALSE)
    } else {
      kable(spcdtab())
      }
```
:::

::: {}
#### Damage Agent Codes / Names
```{r, include=T, echo=F, warning=FALSE, message = FALSE}
if (input$format == "HTML"){
    datatable(damcdtab(), rownames= FALSE)
    } else {
      kable(damcdtab())
      }
```
:::
::::::::::::::


## General Notes / Assumptions

`r htmltopdf(assumptext())`

## References

`r htmltopdf(reftext())`





